\documentclass[a4paper]{article}
\topmargin	-10mm
\oddsidemargin	0mm
\textwidth	150mm
\textheight	230mm

\input{commands}

\begin{document}

%----------------------------------------------------------------------
\title{Towards Ocean Model-Driven Robotic Exploration
  \\\vspace{5mm}
 Authors response 
  \\\vspace{5mm}
% \small{Manuscript ROB-17-0148}
}


\date{\today}

\maketitle

The authors are grateful to the reviewers for their insightful and
constructive comments and suggestions. A detailed response to the
review follows.

The principal revisions made to the manuscript are:

\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

\item \kc{To be filled once the following below is completed}

\end{itemize}

Note that the reference to page $X$, line $Y$, are made relative to the
previously submitted manuscript.

%======================================================================
\section*{Answers to Reviewer 1}
%======================================================================
All changes requested from Reviewer 1 are marked as \rone{blue text} in the updated manuscript. \setcounter{reviewer}{1} \\

%----------------------------------------------------------------------


\setcounter{reviewer}{1}

\begin{answers}

\item {\it Major comment:}

  \reply: Abstract should be re-written to reflect the findings of the
  study. The literature review can be extended (e.g. Ford et al. 2022)
  in the introduction section. In the methodology, the authors challenge
  the fully dynamical ocean models with a geostatical model. I
  understand the constraints to use a dynamical model. However, the
  authors should carefully demonstrate various aspects of the
  statistical model. They estimate the uncertainty via an ensemble. How
  is this ensemble generated? How reliable is it? Can we see a
  temperature prediction of the geostatistical model to be compared with
  CMEMS model to investigate where are the major differences?

  \begin{enumerate}

  \item {\it Literature review}

  \action: 
  Thank you for pointing out the Ford et.al
  paper \cite{ford22}. Their work indeed is synergistic with ours,
  caveat, we use powered AUVs, not gliders as noted already in the
  Introduction; that is AUVs which can articulate their paths in ways
  gliders cannot by virtue of a propeller and their software control
  system. We have modified the text to read:

  \begin{quote}

    \cite{lermusiaux07} and \cite{ford22} explore such a concept, the
    former in simulation, the latter focusing on chlorophyll
    while we targeted temperature gradients; moreover the novelty of our
    work is in
  demonstrating the applicability of using our AUVs that can control
  their trajectory using onboard computation. Our setup is also simpler
  with CMEMS seeding the stochastic model augmenting the need to show
this
  \emph{sample-assimilate-predict-direct} loop in a cost effective
  manner relying on small boats, the use of multiple AUVs while
  relying on our mature technology suite for command/control of
  vehicles.

\end{quote}

\item {\it Methodology} \kc{- LEONARDO (LA) and FILIPA (AFD) - check how
  this was explained on the previous paper and how we can introduce that
  here in a very simple manner.}

\end{enumerate}
  
\item {\it Specific Comments:}

  \begin{enumerate}

  \item \reply: L65: This paragraph may be better placed in the
    abstract or
  conclusions to serve as a summary of the work’s contribution.

  \action: We have modified the abstract per the suggestion while
  keeping the text as is in the body of the article, by adding the
  following to the abstract:

  \begin{quote}
    
    The novelty of this effort is
  threefold; the demonstration of coupling models with AUV sampling,
  the importance of targeted sampling and the impact of loop closure
  towards model prediction.
  \end{quote}

\item \reply: L100/L105: For better geographical context, please mark
  the locations (Berlengas archipelago, Obidos lagoon, etc.) on Figure 1
  for readers unfamiliar with the region.

  \action: \kc{RM will take care}

\item \reply: L116: This paragraph appears to be more appropriate for
  the Methods section, as it describes the system’s operational aspects.

  \action: We have placed the entire paragraph at the end of the Methods
  section as suggested.

\item \reply: L118: Please clarify the phrase “to democritize ocean
  model prediction.” Does this imply providing more cost-effective
  prediction systems?

\action: This has been clarified with the following text, now at the end
of the Methods section. 


\begin{quote}
  Our end goal is to democratize ocean model prediction, using low-cost
  operations
  integrated with open-source software to enable spatio-temporal data
  assimilation from mobile or immobile robots across space, \ldots
  \end{quote}

\item \reply: L129: Detail the specific methodology used for the
  uncertainty estimate.
  
  \action: \kc{LA and AFD - related to General Comments}

\item \reply: L147: The correct acronym for the E.U. Copernicus Marine
  Service is CMEMS, not CMS; please correct this throughout the text.
  Furthermore, CMEMS is a service, not a model. Please adapt the text to
  refer to the specific model (e.g., NEMO, ROMS, or HYCOM) and the
  system name (e.g., IBI-MFC system), and include explicit citations to
  their scientific publications.

  \action: Our effort leveraged the NEMO model within CMEMS using the
  IB-MFC domain. This has
  now been clarified in the Methods section including citations as
  suggested with the following text: 

  \begin{quote}

    Our methodology relies on ensembles of geostatistical realizations,
    each representing a state of an ocean variable field conditioned to
    \emph{a priori} by the deterministic CMEMS NEMO model
  \cite{madec97,madec15} using the Iberian Biscay Irish domain. We use
  models \dots
    \end{quote}

    Note that the reference to CMS has been replaced manuscript--wide
    with CMEMS.

\item \reply: L157: Please provide a figure showing the temperature
  fields after the geostatistical simulation.
  \action: \kc{AFD will take care}

\item \reply: L166: Elaborate on the process used to generate the
  ensemble of geostatistical realizations.

  \action: \kc{AFD - we need to be careful here because the paper is already
  super long.}

\item \reply: L173: The description of “sequential assimilation” is
  unclear. Please specify the exact scheme used (e.g., Ensemble Kalman
  Filter) and clarify whether differences with observations are
  calculated before or after assimilation.

  \action: \kc{LA and AFD will take care}

\item \reply: L192: The statement on spatial propagation being
  horizontal with implicit vertical consistency is not immediately
  intuitive. A figure illustrating this concept would be very helpful to
  the reader.

  \action: \kc{ and AFD - I think you have this figure in the previous
  paper. It will be impossible to add one more figure, so we will cite
  it. Make sure that I'm thinking right.}

\item \reply: L204: Please provide a demonstration or figure of the
  “uncertainty map provided by the model.”
  
  \action: \kc{LA, AFD and LB - I think we have this both in Lucrecia and
  Filipa's papers}

\item \reply: L231: The text mentions that the algorithm ensures
  “vehicle endurance and safety constraints.” Please describe how you
  constrain the algorithm for safety.

  \action: \kc{RM and JBS will take care}

\end{enumerate}


\item{\it Minor corrections:} 

  \begin{enumerate}
    
  \item
    \reply: L16: Correct “worlds oceans” to world’s ocean.
    
    \action: Fixed.

  \item \reply: L47: Please be more precise with a quantitative
    assessment to define “lower effective speeds.”
    
    \action: L46 refers to the speeds at which our powered AUVs operate.
    This sentence was meant to provide the comparision with gliders
    which typically operate at lower speeds. We do not believe any
    modification is necessary. 

  \item \reply: L56: Explain the meaning of onboard computation.
    
    \action: Our (and most) AUVs, have far more computational capacity
    on board, akin to a standard laptop; this is not the case with the
    current generation of gliders. For instance, our AUVs carry a Intel
    Core i7-1065G7 CPU rated at 1.30GHz running the Linux operating
    system. Advanced methods of embedded reasoning such as in our
    earlier work \cite{das11b,rajan12,chrpa15,fossum18,fossum19,fossum19b} among
    others, allow onboard computing. While embedded reasoning was neither
    the highlight nor used in this work, the vehicles used allow for
    such capability. We have modified the text to read:

    \begin{quote}

      \ldots moreover the novelty of our work is
  in demonstrating the applicability of using our AUVs that can
  control their trajectory using onboard computation. While well
  suited for embedded reasoning for adaptive sampling
  \cite{das11b,rajan12,chrpa15,fossum18,fossum19,fossum19b}, this
  effort used shore-generated waypoints distributed to all AUVs, a
  task which could be moved onboard in the future. 

    \end{quote}

  \item L112: Correct “physicalbiogeochemical” to
    physical-biogeochemical.
    
    \action: Corrected.

  \item \reply: L285: A panel number for the figure is missing.
    
    \action: Corrected in this and a number of other locations. 
    
  \item \reply L336: Note that XP2 was also operated on the 31st.
    
    \action: Corrected.

  \item \reply: L348: Correct the trajectory labels from C14 to C1-4 and
    D14 to D1-4. Fig. 4: Provide more details on the sea surface
    temperature (SST) datasets in the caption (e.g., CMEMS satellite SST L3?). Fig.
    5: The colorbar description (“Shallower layers are shown in warmer
    colors…”) seems to contradict the visual representation. Please
    verify the description against the colorbar.
    
    \action: LB check the colors

  \item \reply: L362 / Fig. 6: A panel code is missing for Figure 6.
    Please also add an x-axis label to panel (c) of Figure 6.
    
    \action: Panel references are fixed. 

    \item \reply: L430: Correct the compound word to
      upwelling-relaxation.

      \action: Corrected
      
  \item \reply: L432: Correct the compound word to data collected.
    
    \action: Corrected the use of hyphens to reflect the following text:

    \begin{quote}

      This transition can explain why assimilation of southern
      data, collected under a single dynamical regime, fails to improve
      and in fact, degrades forecasts.

      \end{quote}

  \item \reply: L433: Correct the compound word to regime fails.

    \action: Corrected as noted above.

  \end{enumerate}
  
\end{answers}


% \newpage
%======================================================================
\section*{Answers to Reviewer 2}
%======================================================================
All changes requested from Reviewer 2 are marked as \rtwo{red text}
in the updated manuscript. \\


\setcounter{reviewer}{2}

\begin{answers}

\item {\it{General Comments:}}


\item {\it The paper is generally well constructed, well written and accurate. I appreciate the authors’ efforts. For real-time ocean awareness, glider missions are part of a sample-assimilate predict-direct loop. Here, the authors review an experiment where they demonstrate an ability to send gliders on sampling missions, assimilate the data, predict the next day with a low computational cost forecast model, and then generate guidance based on the forecast. All of these components have been developed by the authors and their collaborators. Bring all these components together represents a significant accomplishment. Some parts of their work are novel - the forecast model has low computational cost, and thus appropriate for low-powered/limited-resource environments. The forecast model can also generate uncertainty maps, which are useful for generating glider guidance. The gliders are of their own design. I also appreciate the discussion of the observed subgridscale variability in section 5. Some parts of the paper would benefit additional attention. There is no discussion of the gliders speed through water nor typical water velocities in the region. There is no comparison of predicted glider trajectories versus guidance fields. Also, there is no comparison of predicted trajectories with actual glider trajectories. The combination of these things makes it hard for me to evaluate how well the guidance was followed. There is no comparison of the low-cost forecast model versus the global forecast that the low-cost model is based on. So, it makes it hard to say if this approach improves forecast skill over the global model forecast. I think the paper would be strengthed by showing how well the gliders followed their guidance, showing that the observations reduce uncertainty (maybe just put the glider trajectories on the uncertainty maps?), and showing the performance of the low-cost model versus the parent model.} \\ 

\rmcomment{\\
 A - Need to correct the reviewer about gliders vs AUVs\\\\
 B - \textit{gliders speed through water nor typical water velocities in the region} - RM will take care\\\\
 C - \textit{...comparison of predicted glider trajectories versus guidance fields. Also, there is no comparison of predicted trajectories with actual glider trajectories.... I think the paper would be strengthed by showing how well the gliders followed their guidance, showing that the observations reduce uncertainty (maybe just put the glider trajectories on the uncertainty maps?)} - LB will evaluate how to handle this in terms of figures\\\\
 D - \textit{comparison of the low-cost forecast model versus the global forecast that the low-cost model is based on} - LA and AFD will check this. same reply as R1 in the general comments

}

\textbf{Are the figures and references suitable? }
\item{\it The figures are mostly suitable. Figure 6 needs some clarification. The references are suitable. For Figure 6, part (a) is a different size than all the other parts of Figure 6. Additionally, comparing to 6(a) to 6(bc), there is no specification of the data that the forecasts are compared to. For part (c), the units are missing.}

\rmcomment{LA and AFD - check this comment about fig 6}\\

\textbf{Specific Comments}
\item{\it After reviewing the references, it looks like all the components of this system - the low-cost model, the assimilation, the guidance, and the gliders - have been build and integrated by the authors and colleagues. This study looks like a first end-to-end demonstration. If that is true, I think it is okay to strongly and clearly state that in the introduction. That is a significant achievement.}
\rmcomment{True. RM and KR will take care}\\

\item{\it It would be valuable to see maps of the uncertainty and the predicted and actual glider paths on the uncertainty maps - c.f. Figure 4. }
\rmcomment{LB will take care - same question in general comments}\\

\item{\it Figures 5 and 6 demonstrate that assimilating glider data results in lower forecast error than not assimilating glider data. It would be nice to see that the system outperforms CMS. Also, Fig. 6c needs unit labels.  }
\rmcomment{LA and AFD - check this. same question in general comments}\\

\item{\it Lines 410-439: An alternative explanation for assimilation experiments C4 and D4 performing worse than the free run is that the background error horizontal covariance length scales are over large. When the update is made to the background in experiments C4 and D4, how far away is the background corrected? Is that typical of feature size in the region? Fig. 4a suggests eddies of ~15 km in diameter. That said, construction of 
background error covariances remains a rich topic.}\\
\rmcomment{LA and AFD, can you check what our approach should be here. I don't interpret this as a question but more as a comment, but I'm not sure.}\\

\textbf{Minor Comments}
\item{\it At lines 74-79, section 4 is not mentioned}
\rmcomment{RM will take care}\\

\item{\it At line 178: “With AUV sampling denser than the grid resultion 
(approximately 0.0035 deg., corresponding to about 300-400 m at the study 
latitude)…” Unclear to me if this is the model horizontal resolution or the 
typical distance between glider surfacings/profiles. }
\rmcomment{RM will take care}

\item{\it Line 521-523. I don’t understand the motivation to demonstrate the viaibility of low-cost models in the cloud. Once it is solved on a dedicated system, running it in the cloud is a software engineering task. It sounds to me that the benefit of portable, low-cost models is not necessarily running in the cloud but “containerization” - able to running a wide range of computer hardware, independent of the specific hardware details. This includes AUVs, as you point out. I think there are two benefits of cloud computing for ocean forecasting. First, access an ocean model anywhere there is an internet connection and second, acquire varying computer resources 
proportional to compute need (or forecast need), whether it is low cost or 
high cost. I strongly agree with the motivation to look at forecast skill at a range of time horizons and compute costs }
\rmcomment{RM and KR will take care}

\end{answers}
% ======================================================================
 \section*{Author Changes}
% ======================================================================
 The following changes were made by the authors to provide more
 clarity over and above those suggested by the reviewers. The author
 changes are marked as \me{violet text} in the updated manuscript.

\begin{enumerate}

\item 

\begin{quote}

\end{quote} 
To:
\begin{quote}
\end{quote}

\item 

\begin{quote}

\end{quote}

\end{enumerate}


\bibliographystyle{IEEEtran}
\footnotesize{
  \bibliography{../ref}
}

%======================================================================
\end{document}
%======================================================================

